Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-06-11 Master Thesis/synopsis.tex
Update

Assume that $w = w_0|_K$ is not henselian. Then there exists a $w_1\in|\operatorname{Spv}(K_0)|$ with $w_1\neq w_0$ and such that $w = w_1|_K$, for otherwise $w$ would have a unique extension to $K_0$ and hence to $\overline{K}$ since $w_0$ is henselian.

Let $(K_0)_{w_1}$ be a henselization of $K_0$ with respect to $w_1$ and let $K_w$ be a henselization of $K$ with respect to $w$ in $(K_0)_{w_1}$. Then $(K_0)_{w_1}$ is the composite of $K_0$ and $K_w$ and we have $\ell^\infty\nmid [(K_0)_{w_1} : K_w]$.

@@

\begin{remark}\label{3.3}
If $L\subset M$ are intermediate fields of $\overline{K}/K$, which are finite over $K$, then $\sigma$ also induces bijections
\[ \operatorname{Spv}_\text{d}^{(1)}(L) \to \operatorname{Spv}_\text{d}^{(1)}(L'),\qquad \operatorname{Spv}_\text{d}^{(1)}(M) \to \operatorname{Spv}_\text{d}^{(1)}(M') \]
such that the following diagram commutes:
\[\begin{tikzcd}
\operatorname{Spv}_\text{d}^{(1)}(M) \rar["\sigma"]\dar & \operatorname{Spv}_\text{d}^{(1)}(M')\dar\\
\operatorname{Spv}_\text{d}^{(1)}(L) \rar["\sigma"'] & \operatorname{Spv}_\text{d}^{(1)}(L')
\end{tikzcd}\]
where the vertical maps are restrictions. Let $w\in\operatorname{Spv}_\text{d}^{(1)}(M),\ v=w|_L$ and $w'=\sigma(w),\ v'=\sigma(v)$. Then the following diagram commutes:
\[ \begin{tikzcd}
k(w') \rar["\varphi_w"] & k(w)\\
k(v') \rar["\varphi_v"']\uar[hook] & k(v)\uar[hook]
\end{tikzcd} \]
By limit process we see that $\sigma$ induces a bijection $\sigma: \operatorname{Spv}_\text{d}^{(1)}(\overline{K})\to \operatorname{Spv}_\text{d}^{(1)}(\overline{K'})$. Let $\overline{v}\in \operatorname{Spv}_\text{d}^{(1)}(\overline{K})$ and $G_{\overline{v}}\subset G_K$ be the decomposition group of $\overline{v}$. Then we have $\sigma(G_{\overline{v}}) = G_{\sigma(\overline{v})}$. It follows for any $g\in G_K$:
\[G_{\sigma(g\overline{v})} = \sigma(G_{g\overline{v}}) = \sigma(g^{-1} G_{\overline{v}}g) = \sigma(g)^{-1} G_{\sigma(\overline{v})}\sigma(g) = G_{\sigma(g)\sigma(\overline{v})}\]
Therefore, $\sigma(g\overline{v})=\sigma(g)\sigma(\overline{v})$.
\end{remark}

@@

Let $P\in X_0$ and $v_P:K\to\mathbb{Z}\cup\{\infty\}$ be the corresponding normalized discrete valuation and let $U_P$ be the group of units of the valuation ring $\mathcal{O}_P$. We denote the natural projection $U_P\to k(P)^\times,\ x\mapsto x\mod P$ by $r_P$.

@@

\begin{proof}
\begin{enumerate}[(i)]
\item Consider the absolute inertia field $L = K_P^\text{t}$ of $K_P$ and let $w$ be the unique extension of $v_P$ in $L$. By Theorem \ref{theorem-absoluteinertia} (i), $w$ is discrete as well and we have $k(w) = k(v_P)^\text{s}$. Observe the following exact sequence:
\[ 1 \longrightarrow U_w/n \longrightarrow L^\times /n\longrightarrow \mathbb{Z}/n\mathbb{Z}\longrightarrow 0 \]
Since $k(w)$ is separably closed, with $\operatorname{char} k(w)=0$ and Hensel's lemma we can see that $U_w$ is a divisible group, hence we get an isomorphism $L^\times /n\cong \mathbb{Z}/n\mathbb{Z}$ induced by $w$. Therefore, the canonical map $K^\times/n\to L^\times/n$ can be identified with the map $K^\times/n \to\mathbb{Z}/n\mathbb{Z}$ induced by $v_P$. Now the following commutative diagram with exact columns gives us the result:
\[ \begin{tikzcd}
1 \ar[d] & 1 \ar[d]\\
U_{P'}/n \ar[r, dashed]\ar[d] & U_P/n\ar[d]\\
K'^\times /n \ar[r, "\varphi_n"]\ar[d] & K^\times/n\ar[d]\\
L'^\times /n \ar[r, "\varphi_n"'] &  L^\times /n
\end{tikzcd}\]
Note that $(K^\text{t}_P)' = K'^{\text{t}}_{P'}$ as proven in Lemma \ref{3.2}.
\item Consider the field extension $K_P/K$. By Remark \ref{remark-localphi-extension}, without loss of generality, we can assume $K=K_P$. Since $\operatorname{char}k(P)=0$, by Hensel's lemma, $r_P/n$ is an isomorphism and the composition
\[\begin{tikzcd}
k(P)^\times / n\ar[rr, "(r_P/n)^{-1}"] && U_P/n \ar[rr, hook] && K_P^\times/n
\end{tikzcd}\]
coincides with the inflation $\operatorname{inf}: \mathrm{H}^1(k(P), \mu_n) \to \mathrm{H}^1(K_P,\mu_n)$. The same holds for $P'$ due to Lemma \ref{3.2}. Therefore, the assertion follows from the commutativity of the following diagram:
\[ \begin{tikzcd}
k(P')^\times/n \ar[r, "\cong"] \ar[d, "\varphi_P/n"'] & \mathrm{H}^1(k(P'),\mu_n) \ar[r, "\mathrm{inf}"] \ar[d]& \mathrm{H}^1(K'_{P'}, \mu_n) \ar[d, "\varphi_n"] \\
k(P)^\times/n \ar[r, "\cong"'] & \mathrm{H}^1(k(P),\mu_n)  \ar[r, "\mathrm{inf}"'] & \mathrm{H}^1(K_{P}, \mu_n)
\end{tikzcd}  \qedhere\]
\end{enumerate}
\end{proof}

>>>>

For the sake of contradiction, assume that $w = w_0|_K$ is not henselian. Then there exists a $w_1\in|\operatorname{Spv}(K_0)|$ with $w_1\neq w_0$ and such that $w = w_1|_K$, for otherwise $w$ would have a unique extension to $K_0$ and hence to $\overline{K}$ since $w_0$ is henselian.

Let $(K_0)_{w_1}$ be a henselization of $K_0$ with respect to $w_1$ and let $K_w\subset (K_0)_{w_1}$ be a henselization of $K$ with respect to $w$. Then $(K_0)_{w_1}$ is the composite of $K_0$ and $K_w$ and we have $\ell^\infty\nmid [(K_0)_{w_1} : K_w]$.

@@

\begin{remark}\label{3.3}
If $L\subset M$ are intermediate fields of $\overline{K}/K$, which are finite over $K$, then $\sigma$ also induces bijections
\[ \operatorname{Spv}_\text{d}^{(1)}(L) \to \operatorname{Spv}_\text{d}^{(1)}(L'),\qquad \operatorname{Spv}_\text{d}^{(1)}(M) \to \operatorname{Spv}_\text{d}^{(1)}(M') \]
such that the following diagram commutes:
\[\begin{tikzcd}
\operatorname{Spv}_\text{d}^{(1)}(M) \rar["\sigma"]\dar & \operatorname{Spv}_\text{d}^{(1)}(M')\dar\\
\operatorname{Spv}_\text{d}^{(1)}(L) \rar["\sigma"'] & \operatorname{Spv}_\text{d}^{(1)}(L')
\end{tikzcd}\]
where the vertical maps are restrictions. Let $w\in\operatorname{Spv}_\text{d}^{(1)}(M),\ v=w|_L$ and $w'=\sigma(w),\ v'=\sigma(v)$. Then the following diagram commutes:
\[ \begin{tikzcd}
k(w') \rar["\varphi_w"] & k(w)\\
k(v') \rar["\varphi_v"']\uar[hook] & k(v)\uar[hook]
\end{tikzcd} \]
By limit process we see that $\sigma$ induces a bijection $\sigma: \operatorname{Spv}_\text{d}^{(1)}(\overline{K})\to \operatorname{Spv}_\text{d}^{(1)}(\overline{K'})$. Let $\overline{v}\in \operatorname{Spv}_\text{d}^{(1)}(\overline{K})$ and $G_{\overline{v}}\subset G_K$ be the decomposition group of $\overline{v}$. Then we have $\sigma(G_{\overline{v}}) = G_{\sigma(\overline{v})}$. It follows for any $g\in G_K$:
\[G_{\sigma(g\overline{v})} = \sigma(G_{g\overline{v}}) = \sigma(g^{-1} G_{\overline{v}}g) = \sigma(g)^{-1} G_{\sigma(\overline{v})}\sigma(g) = G_{\sigma(g)\sigma(\overline{v})}\]
Therefore, $\sigma(g\overline{v})=\sigma(g)\sigma(\overline{v})$ by Corollary \ref{corollary-fkschmidt}.
\end{remark}

@@

Let $P\in X_0$ and $v_P:K\to\mathbb{Z}\cup\{\infty\}$ be the corresponding normalized discrete valuation and let $U_P$ be the group of units of the valuation ring $\mathcal{O}_P$. We denote the natural projection $U_P\to k(P)^\times,\ x\mapsto x\mod P$ by $r_P$. We shall also write $K_P$ for a henselization of $K$ with respect to $v_P$.

@@

\begin{proof}
\begin{enumerate}[(i)]
\item Consider the absolute inertia field $L = K_P^\text{t}$ of $K_P$ and let $w$ be the unique extension of $v_P$ in $L$. By Theorem \ref{theorem-absoluteinertia} (i), $w$ is discrete as well and we have $k(w) = k(v_P)^\text{s}$. Observe the following exact sequence:
\[ 1 \longrightarrow U_w/n \longrightarrow L^\times /n\longrightarrow \mathbb{Z}/n\mathbb{Z}\longrightarrow 0 \]
Since $k(w)$ is separably closed, with $\operatorname{char} k(w)=0$ and Hensel's lemma we can see that $U_w$ is a divisible group, hence we get an isomorphism $L^\times /n\cong \mathbb{Z}/n\mathbb{Z}$ induced by $w$. Therefore, the canonical map $K^\times/n\to L^\times/n$ can be identified with the map $K^\times/n \to\mathbb{Z}/n\mathbb{Z}$ induced by $v_P$. Now the following commutative diagram with exact columns gives us the result:
\[ \begin{tikzcd}
1 \ar[d] & 1 \ar[d]\\
U_{P'}/n \ar[r, dashed]\ar[d] & U_P/n\ar[d]\\
K'^\times /n \ar[r, "\varphi_n"]\ar[d] & K^\times/n\ar[d]\\
L'^\times /n \ar[r, "\varphi_n"'] &  L^\times /n
\end{tikzcd}\]
Note that $(K^\text{t}_P)' = K'^{\text{t}}_{P'}$ as proven in Lemma \ref{3.2}.
\item Consider the field extension $K_P/K$. By Remark \ref{remark-localphi-extension}, without loss of generality, we can assume $K=K_P$. Since $\operatorname{char}k(P)=0$, by Hensel's lemma the map $r_P/n: U_P/n\to k(P)^\times/n$ is an isomorphism and the composition
\[\begin{tikzcd}
k(P)^\times / n\ar[rr, "(r_P/n)^{-1}"] && U_P/n \ar[rr, hook] && K_P^\times/n
\end{tikzcd}\]
coincides with the inflation $\operatorname{inf}: \mathrm{H}^1(k(P), \mu_n) \to \mathrm{H}^1(K_P,\mu_n)$. The same holds for $P'$ due to Lemma \ref{3.2}. Therefore, the assertion follows from the commutativity of the following diagram:
\[ \begin{tikzcd}
k(P')^\times/n \ar[r, "\cong"] \ar[d, "\varphi_P/n"'] & \mathrm{H}^1(k(P'),\mu_n) \ar[r, "\mathrm{inf}"] \ar[d]& \mathrm{H}^1(K'_{P'}, \mu_n) \ar[d, "\varphi_n"] \\
k(P)^\times/n \ar[r, "\cong"'] & \mathrm{H}^1(k(P),\mu_n)  \ar[r, "\mathrm{inf}"'] & \mathrm{H}^1(K_{P}, \mu_n)
\end{tikzcd}  \qedhere\]
\end{enumerate}
\end{proof}
